---
title: "p8105_hw2_jp3665"
author: "Joana Petrescu"
date: "9/30/2020"
output: github_document
---

```{r, message = FALSE}

library(readxl)
library(dplyr)
library(tidyverse)

```


# Problem 1

## Read the Mr. Trash Wheel dataset

I already created a subdirectory for the data used in this problem set and I downloaded the `Trash-Wheel-Collection-Totals-8-6-19.xlsx` file to this subdirectory

Read each sheet in the Excel file and save each sheet as a variable

```{r}

excel_sheets("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx")

Mr_data <- read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "Mr. Trash Wheel", skip = 1)

Prof_data <- read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "Professor Trash Wheel", skip = 1)

Capt_data <- read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "Captain Trash Wheel", skip = 1)

precip_2019_data <- read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "2019 Precipitation", skip = 1)

precip_2018_data <- read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "2018 Precipitation", skip = 1)

precip_2017_data <- read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "2017 Precipitation", skip = 1)

precip_2016_data <- read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "2016 Precipitation", skip = 1)

precip_2015_data <- read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "2015 Precipitation", skip = 1)

precip_2014_data <- read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "2014 Precipitation", skip = 1)


```

## Clean up Mr. Trash Wheel dataset

Remove blank columns from`Mr_data`

```{r}

head(Mr_data)

Mr_data <- select(Mr_data, -...15, -...16, -...17)

```

Omit rows that do not include dumpster specific names

```{r}

Mr_data <- Mr_data %>% drop_na(Dumpster)

Prof_data <- Prof_data %>% drop_na(Dumpster)

Capt_data <- Capt_data %>% drop_na(Dumpster)

```

Round sports balls to nearest integer and convert to integer

```{r}

Mr_data$`Sports Balls` <-  round(Mr_data$`Sports Balls`, digits = 0)

Mr_data$`Sports Balls` <-  as.integer(Mr_data$`Sports Balls`, digits = 0)

```


## Clean up precipitation data

Clean and combine precipitation data for 2017 and 2018

```{r}

precip_2017_data <- precip_2017_data %>% drop_na(Month)

precip_2018_data <- precip_2018_data %>% drop_na(Month)

precip_2017_data <- mutate(precip_2017_data, Year = 2017)

precip_2018_data <- mutate(precip_2018_data, Year = 2018)

months <- list(name = month.name, num = 1:12)

precip_2017_data <- precip_2017_data %>% mutate(Month = replace(Month, months$num, months$name))

precip_2018_data <- precip_2018_data %>% mutate(Month = replace(Month, months$num, months$name))

precip_2017_2018_data <- rbind(precip_2017_data, precip_2018_data)

```


## Description of Mr. Trash Wheel dataset

This dataset now contains 344 observations of 14 variables. The variables represent the _dumpster number_, _date_, and number of the following objects that were found in each dumpster: _plastic bottles_, _polystyrene objects_, _cigarette butts_, _glass bottles_, _grocery bags_, _chip bags_, and _sports balls._ The values for _sports balls_ are currently in integer form.

```{r}

Mr_data_2017 <- filter(Mr_data, Year == "2017")

median(Mr_data_2017$`Sports Balls`)

```

The median number of sports balls found in each dumpster in 2017 is 8.

## Description of precipitation dataset

The precipitation dataset for 2017 and 2018 now contains 24 observations for three variables: the _year_, _month_, and _total precipitation_. The month column now contains full month names.

```{r}

sum(precip_2018_data$Total)

```

The total precipitation in 2018 was 70.33.

# Problem 2

## Read NYC Transit Subway Data

I already created a subdirectory for the data used in this problem set and I downloaded the `NYC_Transit_Subway_Entrance_And_Exit_Data.csv` file to this subdirectory.

Read the `.csv` file

```{r}

NYC_subway_dataset <- read_csv(file = "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv")

```
Retain line, station name, station latitude/longitude, routes served, entry, vending, entrance type, and ADA compliance

```{r}

NYC_subway_data <- select(NYC_subway_dataset, "Line", "Station Name", "Station Latitude", "Station Longitude", "Route1", "Route2", "Route3", "Route4", "Route5", "Route6", "Route7", "Route8", "Route9", "Route10", "Route11", "Entry", "Vending", "Entrance Type", "ADA")

```

Convert entry variable from _yes_ or _no_ values to logical variable

```{r}

NYC_subway_data$Entry <- ifelse(NYC_subway_data$Entry == "YES", TRUE, FALSE)

```


## Description of NYC Transit Subway Data

This dataset describes subway stations in NYC. The dataset now consists of 1868 observations for 18 different varaibles. I omited columns from the original data set, retaining only variables for the _subway line_, _station name_, _station latitude and longitude_, _routes_, _station entry_, _vending_, _entrance type_, and _ADA compliance_. I then converted the variable describing whether or not entrance is allowed, _entry_, from a character variable (yes or no) to a logical variable (`TRUE` or `FALSE`). These data are now tidy in that they are more suited to answering the questions that follow about subway station entrances.

Identify distinct subway stations

```{r}

NYC_subway_stations <- select(NYC_subway_data, "Line", "Station Name")

nrow(distinct(NYC_subway_stations))

```

There are 465 distinct subway stations based on distinct combinations of subway lines and station names provided.

Number of ADA compliant subway stations

```{r}

nrow(filter(NYC_subway_data, ADA == TRUE) %>% select("Line", "Station Name") %>% distinct())

```

There are 84 distinct subway stations that are ADA compliant.

```{r}

no_vending_data <- filter(NYC_subway_data, Vending == "NO")

no_vending_entry_data <- filter(no_vending_data, Entry == TRUE)

nrow(no_vending_entry_data) / nrow(no_vending_data)

```

Approximately 38% of station entrances/exits with vending allow entrance based on this data.

## Reformat NYC Transit Subway data by route name

Reformat data so that route number and route name are distinct variables

```{r}

NYC_subway_data <- unite(NYC_subway_data, "Route Name", Route1:Route11, sep = "", na.rm = TRUE)

NYC_subway_data <- mutate(NYC_subway_data, "Route Number" = str_count(NYC_subway_data$`Route Name`))

NYC_subway_data <- select(NYC_subway_data, "Line", "Station Name", "Station Latitude", "Station Longitude", "Route Name", "Route Number", "Entry", "Vending", "Entrance Type", "ADA")

head(NYC_subway_data)

```

Subset stations that serve A train

```{r}

A_train <- filter(NYC_subway_data, grepl("A", NYC_subway_data$`Route Name`) == TRUE)

nrow(A_train %>% select("Line", "Station Name") %>% distinct())

nrow(A_train %>% filter(ADA == TRUE) %>% select("Line", "Station Name") %>% distinct())

```

There are 60 distinct stations that serve the A train, and 17 of these stations are ADA compliant.


# Problem 3

I already created a subdirectory for the data used in this problem set and I downloaded the FiveThirtyEight data to this subdirectory.

## Read FiveThirtyEight data

Read `pols-month.csv`

```{r}

pols_month_data <- read.csv("./data/fivethirtyeight_datasets/pols-month.csv")

```

Read `unemployment.csv`

```{r}

unemployment_data <- read.csv("./data/fivethirtyeight_datasets/unemployment.csv")

```

Read `snp.csv`

```{r}

snp_data <- read.csv("./data/fivethirtyeight_datasets/snp.csv")

```

## Clean pols_month_data

Clean `pols_month_data` by parsing date into `year`, `month` and `day`, changing numeric months to full month names, and removing `day` column

```{r}

head(pols_month_data)

pols_month_data <- separate(pols_month_data, "mon", c("year", "month", "day"), sep = "-", remove = TRUE, extra = "warn", fill = "warn")

pols_month_data$month <- as.numeric(pols_month_data$month)

months <- list(name = month.name, num = 1:12)

pols_month_data <- pols_month_data %>% mutate(month = replace(month, months$num, months$name))

pols_month_data <- pols_month_data %>% select(-day)

```

Remove `prez_dem` and `prez_gop` and combine into `president` variable

```{r}

pols_month_data <- pols_month_data %>% mutate("president" = ifelse(prez_gop == 1, "gop", "dem"))

pols_month_data <- pols_month_data %>% select(-prez_gop, -prez_dem)

head(pols_month_data)

```

## Clean snp_data

Clean `snp.csv` by parsing `date` into `day`, `month` and `year`, replacing numeric months with full month names, removing the day column, and reorganizing with year and month as the first columns

```{r}

head(snp_data)

snp_data <- separate(snp_data, "date", c("day", "month", "year"), sep = "/", remove = TRUE, extra = "warn", fill = "warn")

snp_data$month <- as.numeric(snp_data$month)

snp_data <- snp_data %>% mutate(month = replace(month, months$num, months$name))

snp_data <- snp_data %>% select(-day)

snp_data <- snp_data %>% select(year, month, close)

head(snp_data)

```


## Clean unemployment_data

Clean unemployment data by changing from wide form to long form and formating first two columns to match other datasets

```{r}

head(unemployment_data)

unemployment_data <- unemployment_data %>% gather("month", "unemployment", -Year)

colnames(unemployment_data) [1] <- "year"

head(unemployment_data)

months_abbrev <- list("name" = month.name, "abb" = month.abb)

unemployment_data$month <- unemployment_data$month %>% str_replace_all("Jan", "January")
unemployment_data$month <- unemployment_data$month %>% str_replace_all("Feb", "February")
unemployment_data$month <- unemployment_data$month %>% str_replace_all("Mar", "March")
unemployment_data$month <- unemployment_data$month %>% str_replace_all("Apr", "April")
unemployment_data$month <- unemployment_data$month %>% str_replace_all("Jun", "June")
unemployment_data$month <- unemployment_data$month %>% str_replace_all("Jul", "July")
unemployment_data$month <- unemployment_data$month %>% str_replace_all("Aug", "August")
unemployment_data$month <- unemployment_data$month %>% str_replace_all("Sep", "September")
unemployment_data$month <- unemployment_data$month %>% str_replace_all("Oct", "October")
unemployment_data$month <- unemployment_data$month %>% str_replace_all("Nov", "November")
unemployment_data$month <- unemployment_data$month %>% str_replace_all("Dec", "December")

```


